Generating Correlated Data for Omics Simulation

### 2024-09-19 manuscript/paper.qmd
Update for wishart method

Simulated data captures the genes' mean and variance accurately (@fig-compare-to-real a-b).
Moreover, we computed the gene-gene correlation on a random subset of `r N_rows_gene_corr` high-expressed genes.
Since only 12 samples are in each real or simulated data set, correlation estimates are very noisy.
However, the distribution of correlations across all pairs of these genes is much closer to that of real data when simulated with dependence than without (@fig-compare-to-real c).

@@

```{r}
#| label: fig-compare-to-real
#| fig-cap: !expr glue::glue("Comparison to real data run on a mouse liver example from GSE77221. (a-b) Comparison of gene (a) mean expression and (b) variance, log-scaled. The line of equality is marked in black. Points are colored according to the density of points in their region. (c) Quantile-quantile plot comparing correlation values of gene pairs from real data and simulated data (both with and without dependence). A random sample of {N_rows_gene_corr} genes with at least {HIGH_EXPR_CUTOFF} reads was used. Values on the diagonal line indicate a match between the simulated and real data sets. (d) Projections onto the top two principal components of the real data set for both real and simulated data. All 8 simulations (96 samples total for each of dependent and independent simulations) shown. (e) Principal component analysis was performed on all data sets and the variance captured by the top components is shown. Unlike (d), these components were fit from each data set considered separately instead of reusing the weights from the real data.")
#| fig-width: 10
#| fig-height: 10
compare_to_real_plot
```

>>>>

Simulated data captures the genes' mean and variance accurately (@fig-compare-to-real a-b).
Moreover, we computed the gene-gene correlation on pairs of high-expressed genes.
Since only 12 samples are in each real or simulated data set, correlation estimates are very noisy.
However, the distribution of correlations across all pairs of these genes is much closer to that of real data when simulated with dependence than without (@fig-compare-to-real c).

@@

```{r}
#| label: fig-compare-to-real
#| fig-cap: !expr glue::glue("Comparison to real data run on a mouse liver example from GSE77221. (a-b) Comparison of gene (a) mean expression and (b) variance, log-scaled. The line of equality is marked in black. Points are colored according to the density of points in their region. (c) Quantile-quantile plot comparing correlation values of gene pairs from real data and simulated data (both with and without dependence). Genes with at least {HIGH_EXPR_CUTOFF} reads were used. Values on the diagonal line indicate a match between the simulated and real data sets. (d) Projections onto the top two principal components of the real data set for both real and simulated data. All 8 simulations (96 samples total for each of dependent and independent simulations) shown. (e) Principal component analysis was performed on all data sets and the variance captured by the top components is shown. Unlike (d), these components were fit from each data set considered separately instead of reusing the weights from the real data.")
#| fig-width: 10
#| fig-height: 10
compare_to_real_plot
```
